---
title: "Decision Tree using Caret Package"
author: "Kumar Rahul"
date: "9 September 2016"
output:
  html_document:
    df_print: paged
---

## In this exercise, we will use the HR dataset and understand the following using caret package:

> 1. Building the decision tree model
2. Creating the Confusion Matrix and ROC plot on train data
3. Creating the Confusion Matrix and ROC plot on test data

There are bugs/missing code in the entire exercise. The participants are expected to work upon them.
***
***

## Here are some useful links:

> 1. **[Read](http://www.ats.ucla.edu/stat/mult_pkg/faq/general/dummy.htm)** about interaction variable coding
2. Refer **[link](http://www.statmethods.net/input/valuelabels.html)** to know about adding lables to factors
3. Refer **[link](http://stackoverflow.com/questions/2342472/recode-relevel-data-frame-factors-with-different-levels)** to relvel factor variables
4. **[Read](http://stats.stackexchange.com/questions/88485/variable-is-significant-through-stepwise-regression-but-not-in-final-models-sum)** about the issues in stepwise regression
5. **[Read](http://topepo.github.io/caret/training.html)** about the modelling activity via caret package
6. The **[complete](http://topepo.github.io/caret/available-models.html)** list of tuning parameter for different models in caret package


***

# Code starts here
We are going to use below mentioned libraries for demonstrating logistic regression:

```{r libraries, echo=TRUE, message=FALSE, warning=FALSE}
library(caret)    #for data partition. Model building
#library(Deducer) #for ROC plot
library(ROCR)     #for ROC plot (other way)
#library(rattle)   #for plotting tree
library(rpart)
```

## Data Import and Manipulation

### 1. Importing a data set

_Give the correct path to the data_

```{r readData, echo=TRUE,tidy=TRUE}
raw_df <- read.csv("/Users/Rahul/Documents/Datasets/IMB533_HR_Data_No_Missing_Value.csv", header = TRUE,sep = ",",na.strings = c(""," ", "NA"))
```

Note that `echo = FALSE` parameter prevents printing the R code that generated the
plot.

### 2. Structure and Summary of the dataset

```{r summarizeData, echo=TRUE,tidy=TRUE}
str(raw_df)
summary(raw_df)
```

Create a new data frame and store the raw data copy. This is being done to have a copy of the raw data intact for further manipulation if needed.

```{r createDataCopy, echo=TRUE,tidy=TRUE}
filter_df <- na.omit(raw_df) # listwise deletion of missing
```

### 3. Create train and test dataset

#### Reserve 80% for **_training_** and 20% of **_test_**

_Correct the error in the below code chunk_

```{r createDataPartition, echo=TRUE,tidy=TRUE}
set.seed(2341)
trainIndex <- createDataPartition(filter_df$Status, p = 0.80, list = FALSE)
train_df <- filter_df[trainIndex,]
test_df <- filter_df[-trainIndex,]
```

We can pull the specific attribute needed to build the model is another data frame. This agian is more of a hygine practice to not touch the **train** and **test** data set directly.

```{r variableUsedinTraining, echo=TRUE,tidy=TRUE}
dt_train_df <- as.data.frame(train_df[,c("DOJ.Extended",
                                             "Duration.to.accept.offer",
                                             "Notice.period",
                                             "Offered.band",
                                             "Percent.difference.CTC",
                                             "Joining.Bonus",
                                             "Gender",
                                             "Candidate.Source",
                                             "Rex.in.Yrs",
                                             "LOB",
                                             "Location",
                                             "Age",
                                             "Status"
)])
```

```{r variableUsedinTesting, echo=TRUE, tidy=TRUE}
dt_test_data <- as.data.frame(test_df[,c("DOJ.Extended",
                                           "Duration.to.accept.offer",
                                           "Notice.period",
                                           "Offered.band",
                                           "Percent.difference.CTC",
                                           "Joining.Bonus",
                                           "Gender",
                                           "Candidate.Source",
                                           "Rex.in.Yrs",
                                           "LOB",
                                           "Location",
                                           "Age",
                                           "Status"
)])
```

***

## Model Building: Using the **caret()** package

The below chunk of code is standarized way of building model using caret package. Setting in the control parameters for the model.

```{r caretControl, echo=TRUE}
objControl <- trainControl(method = "cv", number = 2,
                           summaryFunction = twoClassSummary,
                           classProbs = TRUE,
                           sampling = 'down',
                           savePredictions = TRUE)
```

Using search grid to fine tune the model

```{r echo=TRUE, message=FALSE, warning=FALSE}
search_grid <- expand.grid(cp=seq(0.001,0.02, 0.001))
```

The model building starts here.
> 1. **metric= "ROC"** uses ROC curve to select the best model.Accuracy, Kappa are other options. To use this change twoClassSummary to defaultSummary in **ObjControl**
2. **verbose = FALSE**: does not show the processing output on console

The factor names at times may not be consistent. R may expect **"Not.Joined"** but the actual level may be **"Not Joined"** This is corrected by using **make.names()** function to give syntactically valid names. Type ?rpart.control in console to get the list of parameters which control the tree growth.

```{r caretModel, echo=TRUE, message=FALSE, warning=FALSE}
#dt_train_df$StatusFactor <- as.factor(ifelse(dt_train_df$Status == "Joined", 1,0))
set.seed(766)
levels(dt_train_df$Status) <- make.names(levels(factor(dt_train_df$Status)))
formula <- as.formula(Status~.)

dt_caret_model <- caret:::train.formula(formula,
                     dt_train_df,
                     method = 'rpart',
                     minsplit = 100, 
                     maxdepth = 4,
                     metric = "ROC",
                     tuneLength = 10,
                     trControl = objControl
                     #tuneGrid = search_grid
                     )

dt_caret_model
```


Business rules from the tree.

```{r caretVarImp, echo=TRUE}
dt_caret_model$finalModel
```

Visualize the rules.

```{r}
library(rpart.plot)
rpart.plot(dt_caret_model$finalModel, type = 4,extra = 104)
```

## Model Evaluation

### 1. The prediction and confusion Matrix on train data.

The syntax for prediction in caret is almost similar expect the the **type** attribute expects input as **'raw'** or **'prob'**. In case of prob, the predicted value holds the probability of both positive and negative class.

```{r caretPrediction, echo=TRUE}
#Missing code. May result in error
levels(dt_train_df$Status) <- make.names(levels(factor(dt_train_df$Status)))
caretPredictedClass <- predict(object = dt_caret_model, dt_train_df[,1:12], type = 'raw')
confusionMatrix(caretPredictedClass,dt_train_df$Status)
```

### 2. Confusion Matrix on the test data

The **predict** function is used to get the predicted class on the new dataset.

```{r modelValidation, echo=FALSE,tidy=TRUE}
levels(dt_test_data$Status) <- make.names(levels(factor(dt_test_data$Status)))
dtCaretTestPredictedClass = predict(dt_caret_model, dt_test_data, type = "raw")
confusionMatrix(dtCaretTestPredictedClass,dt_test_data$Status)
```

### 3. ROC Plot on the test data

ROCR package can be used to evaluate the model performace on the test data. The same package can also be used to get the model performace on the test data.

```{r validationROC, echo=FALSE,tidy=TRUE}
#error in below line
dtCaretTestPredictedProbability = predict(dt_caret_model, dt_test_data, type = "prob")
dtPredObj <- prediction(dtCaretTestPredictedProbability[2],dt_test_data$Status)
dtPerfObj <- performance(dtPredObj, "tpr","fpr")
#dev.off()
plot(dtPerfObj,main = "ROC Curve",col = 2,lwd = 2)
abline(a = 0,b = 1,lwd = 2,lty = 3,col = "black")
performance(dtPredObj, "auc")
```

#### End of Document
***
***
