---
title: "Manipal Hospital Model"
author: "Kumar Rahul"
date: "07 January 2017"
output: word_document
---
## In this exercise, we will use the hospital survey dataset (7126 observations across 51 variables) and understand the following:

> 1. Reading the csv file in R
2. Creating data partition
3. Applying knnImpute algorithm to impute the missing value
4. Converting the standarized variables to orginal values

***

```{r, include=FALSE}
library(caret)            #for split and model accuracy
library(RANN)             #for KKNImpute
```

## Preparing data

####Read data from a specified location
```{r, echo=TRUE}
raw.data <- read.csv("/Users/Rahul/Documents/Rahul Office/IIMB/Projects/Manipal Hospital/Combined_Data_Numeric_Score_4212017.csv",head=TRUE,na.strings=c("", " ","#N/A", "NA"), sep=",")
```

####Define an 70%/30% train/test split of the dataset
```{r}
set.seed(4121)
trainIndex <- createDataPartition(raw.data$NPS_Status, p = 0.70, list=FALSE)
data.train <- raw.data[ trainIndex,]
data.test <- raw.data[-trainIndex,]
#write.csv(na.omit(data.test),"Manipal_Hospital_TestSet_04232017.csv")
```

#Pre-Process for Imputation
preprocessing with knnImpute silently removes any columns containing NA values. As a guideline, **K** should be square root of the number of complete cases.
```{r}
#data.train <- data.train[-1]
preProcValues <- preProcess(data.train[,2:51], method = c("knnImpute"),
                            k = 28,
                            knnSummary = mean)
impute.train <- predict(preProcValues, data.train,na.action = na.pass)

#To get the original data
AgeYrs <- impute.train$AgeYrs*preProcValues$std[1]+preProcValues$mean[1]

Estimatedcost <- impute.train$Estimatedcost *preProcValues$std[2]+preProcValues$mean[2]

LengthofStay <- impute.train$LengthofStay *preProcValues$std[38]+preProcValues$mean[38]

CE_NPS <- impute.train$CE_NPS *preProcValues$std[39]+preProcValues$mean[39]

survey.list <- lapply(12:46, function(x) ceiling(impute.train[x]*preProcValues$std[x-9]+preProcValues$mean[x-9]))
imputed.data <- data.frame(data.train[,1:2],
                           AgeYrs,
                           data.train[,4:6],
                           Estimatedcost,
                           data.train[,8:11],
                           survey.list,
                           data.train[,47:48],
                           LengthofStay,
                           CE_NPS,
                           data.train[51])
#write.csv(imputed.data,"Manipal_Hospital_TrainingSet_04232017.csv")
```

End of document

***
